---
title: "Verreauxii_tadpole_code"
output: html_document
date: '2023-10-27'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Statistical analysis: 

(1) that Bd prevalence in tadpoles would be lower than in adults from the same breeding sites 

```{r}
install.packages("pacman")
pacman::p_load(DHARMa, lme4, rptR, ggplot2, dplyr, readr, tidyverse, patchwork, epitools, cowplot, here)

data <- read_csv(here("data", "clean_data.csv")) #all data 
zoop <- read_csv(here("data","zoop.csv")) #microfauna analysis summary
temp_summary <- read_csv(here("data","temp_summary.csv")) #ibutton data 

```

```{r}
#issue with levels and record_type- need to check the levels before running 
str(data)
data$survey_period <- as.factor(data$survey_period)
data$record_type <- factor(data$record_type, levels = unique(data$record_type))
levels(data$record_type)


glm1 <- glmer(Bd_binary ~ record_type + survey_period + scale(mean_temp) + scale(temp) + scale(count) + scale(species_richness) + (1 | site), family = binomial, data = data)

isSingular(glm1)
summary(glm1)
confint(glm1)

simulationOutput <- simulateResiduals(fittedModel= glm1)
simulationOutput
plot(simulationOutput)

```

Adult Model Only- 

```{r}




```

GLM for body size vs prevalence for tadpoles

```{r}
# Fit a GLM with binomial family to look at tadpole body size effects 

glm_model_body <- glm(Bd_binary ~ length_mm, family = poisson, data = all_data)

# Summarize the model
summary(glm_model_body)

```

##GLM's for microfauna 

(3) there would be a negative relationship between Bd prevalence in tadpoles and the abundance and diversity of microfauna, which can consume Bd zoospores.  
WAS found in the GLMM but not accurate yet. 

```{r}
#load zoop dataframe
zoop <- read_csv("zoop.csv")

##GLM to test abundance and diversity vs elevation 
glm_diversity <- glm(shannon_index ~ elevation_m, data = zoop, family = gaussian())
summary(glm_diversity)
```

```{r}
glm_abundance <- glm(count ~ elevation_m, data = zoop, family = poisson())
summary(glm_abundance)

```


##VISUALISATIONS #### JORDANN CLEAN THIS UP!!!! 
##Dan or Ben dont look below here 

##need to log or sqrt all Bd GE equivalance and prevelance for graphs etc ### 

- look at regression for adult prevelance data by TEMP and ELEVATION???  
- visualise Bd prevelance for adults and tadpoles by elevation - stacked bar plot? (sample size above the columns)
NOW making stacked bar plot with sample sizes, Bd prevelance proportions by elevation, for adults and both tadpole sampling periods 

```
# Create a binary column based on the "Bd" column
df <- df %>% mutate(binary_bd = case_when(
  Bd == "Positive" ~ 1,
  Bd == "Negative" ~ 0,
  TRUE ~ NA_real_
))

# Aggregate the data based on elevation, site, record_type, binary_bd, and Survey_period
aggregated_df <- df %>%
  group_by(elevation_m, site, record_type, Survey_period, binary_bd) %>%
  summarise(count = n(), .groups='drop')

# Calculate the total count per group (elevation, record_type, Survey_period)
total_count_per_group <- aggregated_df %>%
  group_by(elevation_m, record_type, Survey_period) %>%
  summarise(total_count = sum(count), .groups='drop')

# Include the total count back into the aggregated DataFrame
aggregated_df <- left_join(aggregated_df, total_count_per_group, by = c("elevation_m", "record_type", "Survey_period"))

# Calculate the proportion for each group, Adults, TE, TL
aggregated_df$proportion <- aggregated_df$count / aggregated_df$total_count

#
# Create a new column with custom facet labels
aggregated_df$custom_facet <- case_when(
  aggregated_df$record_type == "Adult" & is.na(aggregated_df$Survey_period) ~ "Adult",
  aggregated_df$record_type == "Tadpole" & aggregated_df$Survey_period == 1 ~ "Tadpole Early",
  aggregated_df$record_type == "Tadpole" & aggregated_df$Survey_period == 2 ~ "Tadpole Late",
  TRUE ~ as.character(interaction(aggregated_df$record_type, aggregated_df$Survey_period, lex.order = TRUE))
)

# Convert elevation_m to numeric and sort it
aggregated_df$elevation_m <- as.numeric(as.character(aggregated_df$elevation_m))
aggregated_df <- aggregated_df[order(aggregated_df$elevation_m), ]

# Convert it back to factor with ordered levels
aggregated_df$elevation_m <- factor(aggregated_df$elevation_m, levels = unique(aggregated_df$elevation_m))
  
# Create the plot
prevelance_plot <- ggplot(aggregated_df, aes(x = elevation_m, y = proportion, fill = as.factor(binary_bd))) +
  geom_bar(stat = "identity", position = "stack", width = 0.6) +
  geom_text(aes(label = total_count, y = 1.01), position = position_dodge(width = 0.3), vjust = 0, size = 2) +
  facet_wrap(~ custom_facet, scales = "free_x") +
  scale_x_discrete(expand = c(0.1, 0)) +
  scale_fill_manual(values = c("blue", "orange"), labels = c("Negative", "Positive")) +
  labs(
    title = "Proportion of Bd Infections by Survey and Elevation",
    x = "Elevation (m)",
    y = "Proportion",
    fill = "Bd Status"
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

prevelance_plot

# Save the plot with correct dimensions
ggsave("prevelance_plot.png", plot = prevelance_plot, width = 20, height = 10)

```

(2) that Bd prevalence in tadpoles would decrease as the breeding season progressed and water temperatures increased, potentially creating warm microhabitat conditions where tadpoles could avoid or clear Bd infections. 
GLMM model meant to be the test here... but working that one out

- made overall plot of temps from ibuttons 
- comparison boxplot of time periods 

```
#line plot of temperature over time, grouped by site and coloured by depth category
ggplot(extracted_data, aes(x = date, y = temperature, color = depth_category)) +
  geom_line() +
  labs(x = "Date", y = "Temperature (°C)", color = "Depth Category") +
  facet_wrap(~ site, scales = "free_y", ncol = 2) +  # Facet by site, adjust the number of columns as needed
  theme_minimal()

#make a clear distinction between time periods 

# Calculate a reference date as 4 weeks (28 days) before the maximum date
reference_date <- max(extracted_data$date) - 28*24*3600  # 28 days in seconds

# Create a new column 'time_period' to categorize the dates
extracted_data <- extracted_data %>%
  mutate(time_period = ifelse(date >= reference_date, "72 Hours Before Collection", "72 Hours After Deployment"))

# Create side-by-side boxplots of temperature by elevation category, colored by time period
time_period_plot <- ggplot(extracted_data, aes(x = factor(elevation), y = temperature, fill = factor(time_period))) +
  geom_boxplot(position = "dodge") +
  labs(x = "Elevation", y = "Temperature (°C)", fill = "Time Period") +
  theme_minimal()

time_period_plot

```
- for the above graph need to look at the outlier, and see why some elevations have the after of before data missing- might be the extraction function? 
- box plot shows clear increase in temperature. good way to visualise this? 


TADPOLES AND TEMPS 
- visualise amount of time spent in optimum Bd temperatures for the tadpoles  

```
##tadpole records 

#how many hours are the tadpoles in unfavourable conditions? 
ibutton <- ibutton_data_raw

df_unfavorable <- ibutton %>%
  group_by(site, date) %>%
  mutate(unfavorable_hours = sum(temperature > 25))

merged_data <- merge(ibutton, df_unfavorable, by = c("site", "date"))

#this calculates the number of unfavourable hours at each pond. 
unfavorable_hours_by_pond <- merged_data %>%
  group_by(site) %>%
  summarise(total_unfavorable_hours = sum(unfavorable_hours))

#need to change site to factor, not numeric
ggplot(data = unfavorable_hours_by_pond, aes(x = factor(site), y = total_unfavorable_hours)) +
  geom_bar(stat = "identity") +
  labs(x = "Site", y = "Total unfavorable hours") +
  ggtitle("Total Unfavorable Hours above 25 degrees Celcius") +
  theme(plot.title = element_text(hjust = 0.5)) +
  theme(axis.title = element_text(size = 12), axis.text = element_text(size = 10))

#now for under 4 degrees 
low_unfavorable <- ibutton %>%
  group_by(site, date) %>%
  mutate(unfavorable_hours = sum(temperature < 17))

merged_data2 <- merge(ibutton, low_unfavorable, by = c("site", "date"))

#this calculates the number of unfavourable hours at each pond. 
low_unfavorable_hours_by_pond <- merged_data2 %>%
  group_by(site) %>%
  summarise(total_unfavorable_hours = sum(unfavorable_hours))

#visualise this 
#need to change site to factor, not numeric

ggplot(data = low_unfavorable_hours_by_pond, aes(x = factor(site), y = total_unfavorable_hours)) +
  geom_bar(stat = "identity") +
  labs(x = "Site", y = "Total unfavorable hours") +
  ggtitle("Total Unfavorable Hours below 17 degrees Celcius") +
  theme(plot.title = element_text(hjust = 0.5)) +
  theme(axis.title = element_text(size = 12), axis.text = element_text(size = 10))

## daily min and max temps experienced by tadpoles 
# Read in the iButton data
ibutton_data <- ibutton

# Convert the date/time variable to a POSIXct format
ibutton_data$date <- as.POSIXct(ibutton_data$date, format = "%m/%d/%Y %H:%M")

# Create separate columns for date and time
ibutton_data$date <- as.Date(ibutton_data$date)
ibutton_data$time <- format(ibutton_data$date, "%H:%M:%S")

# Create separate columns for shallow and deep temperatures
ibutton_data1 <- pivot_wider(data = ibutton_data, names_from = site, values_from = temperature)

##too many duplicates- try mean 
# Calculate the daily mean temperatures for each site
daily_mean_temp <- ibutton_data %>%
  group_by(date, site, elevation) %>%
  summarize(mean_temp = mean(temperature, na.rm = TRUE))

# Convert pond to a factor
daily_mean_temp$site <- factor(daily_mean_temp$site)

#plot mean temp 
ggplot(data = daily_mean_temp, aes(x = date, y = mean_temp, color = site)) +
  geom_line() +
  scale_color_discrete(name = "Site") +
  labs(x = "Date", y = "Temperature (°C)") +
  ggtitle("Daily Mean Temperatures by Site") +
  theme(plot.title = element_text(hjust = 0.5), legend.position = "bottom") +
  theme(axis.title = element_text(size = 12), axis.text = element_text(size = 10))
  
#Calculate the daily min and max temperatures for each site
daily_temp_summary <- ibutton_data %>%
  group_by(date, site, elevation) %>%
  summarize(min_temp = min(temperature),
            max_temp = max(temperature))
            
# Convert elevation to a factor
daily_temp_summary$elevation <- factor(daily_temp_summary$elevation)

# Plot the daily min and max temperatures for each site
ggplot(data = daily_temp_summary, aes(x = date, y = max_temp, color = elevation)) +
  geom_line() +
  geom_line(aes(y = min_temp)) +
  scale_color_discrete(name = "Elevation") +
  labs(x = "Date", y = "Temperature (°C)") +
  ggtitle("Daily Min and Max Temperatures by Elevation") +
  theme(plot.title = element_text(hjust = 0.5), legend.position = "bottom") +
  theme(axis.title = element_text(size = 12), axis.text = element_text(size = 10))

##remove the outliers if needed
# Filter out rows with temperatures above 35 degrees
daily_temp_summary_filtered <- daily_temp_summary %>%
  filter(max_temp <= 35, min_temp <= 35)

# Convert elevation to a factor
daily_mean_temp$elevation <- factor(daily_mean_temp$elevation)

# daily mean ggplot
dailymeantemp_plot <- ggplot() +
  geom_line(data = daily_mean_temp, aes(x = date, y = mean_temp, color = elevation)) +
  scale_color_discrete(name = "Site") +
  labs(x = "Date", y = "Temperature (°C)") +
  ggtitle("Daily Mean Temperatures (°C) Over Sampling Period") +
  geom_hline(yintercept = 25, color = "red", linetype = 2, linewidth = 1.0) +
  geom_hline(yintercept = 17, color = "red", linetype = 2, linewidth = 1.0) +
  theme(plot.title = element_text(hjust = 0.5), legend.position = "right") + theme(axis.title = element_text(size = 12), axis.text = element_text(size = 10))

dailymeantemp_plot

##plot the daily min and max temps, with chytrid limits - GREAT FOR PAPER 
#Plot the daily min and max temperatures for each site
minmax_plot <- ggplot(data = daily_temp_summary_filtered, aes(x = date, y = max_temp, color = site)) +
  geom_line(aes(x= date, y = max)) +
  scale_color_discrete(name = "Site") +
  labs(x = "Date", y = "Temperature (°C)") +
  ggtitle("Daily Min and Max Temperatures by Site") +
  geom_hline(yintercept = 25, color = "red", linetype = 2, linewidth = 1.0) +
  geom_hline(yintercept = 4, color = "red", linetype = 2, linewidth = 1.0) +
  theme(plot.title = element_text(hjust = 0.5), legend.position = "bottom") + theme(axis.title = element_text(size = 12), axis.text = element_text(size = 10))

# Filter Bd_records for records with record_type "Tadpole"
tadpole_records <- Bd_records %>%
  filter(record_type == "Tadpole", !is.na(ID))

# Now create the plot using the cleaned data
tadpole_plot <- ggplot(tadpole_records, aes(x = date, y = surface_temp, color = Bd)) +
geom_point()+
  geom_hline(yintercept = 25, color = "red", linetype = 2, linewidth = 1.5) +
  geom_hline(yintercept = 17, color = "red", linetype = 2, linewidth = 1.5) +
  labs(x = "Date", y = "Temperature (°C)") +
  ggtitle("Temperatures of Tadpoles at Time of Capture")

tadpole_plot

# Combine the two plots

library(cowplot)

# Combine plots
# Convert date values to Date format
dailymeantemp_plot$data$date <- as.Date(dailymeantemp_plot$data$date)
tadpole_plot$data$date <- as.Date(tadpole_plot$data$date)

# Set the x-axis limits based on the date range of your data
date_range <- range(dailymeantemp_plot$data$date, tadpole_plot$data$date)

# Create the plots
dailymeantemp_plot <- dailymeantemp_plot +
  xlim(date_range)

tadpole_plot <- tadpole_plot +
  xlim(date_range)

# Combine plots side by side
combined_plot <- plot_grid(dailymeantemp_plot, tadpole_plot, labels = c("A", "B"), ncol = 1)

# Display the combined plot
print(combined_plot)

```

Above graph, seems a little off- I may have cut off the ibutton temps a little later accidentally- we definitely deployed on the same date as tadpoles so seems like a processing error- will fix this. 

```
# Create a bar plot of Shannon index by Elevation and Survey_period
ggplot(zoop_summary_elevation, aes(x = factor(elevation_m), y = mean_shannon, fill = factor(Survey_period))) +
  geom_bar(stat = "identity", position = "dodge", width = 0.7) +
  labs(x = "Elevation (m)", y = "Mean Shannon Index", fill = "Survey Period") +
  facet_wrap(~ Survey_period, ncol = 1, scales = "free_y") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1)) + #rotates the labels 
  scale_x_discrete(labels = function(x) format(as.numeric(x), scientific = FALSE, trim = TRUE)) #reduces the decimal places of elevation
  
#export a table 
install.packages("gt")
library(gt)

zoop_summary %>%
  gt() %>%
  tab_header( title = "Zooplankton Species Richness and Shannon Index Summary")

  
##zooplankton abundance ####
# simplify the columns 
zooplankton1 <- zooplankton %>%
    mutate(Site = as.character(Site)) %>%
  select(c(Family, Order, Site, Survey_period, Count))
 
#order across all sites, in two sampling periods 
ggplot(zooplankton1, aes(x = Site, y = Count, fill = Order)) +
  geom_bar(stat = "identity", position = "dodge", width = 0.7) +
  labs(x = "Site", y = "Abundance", fill = "Order") +
  scale_x_discrete(limits = unique(zooplankton1$Site)) +
  facet_wrap(~ Survey_period, ncol = 1, scales = "free_y") +
  theme(axis.text.y = element_text(angle = 90, hjust = 1))

##making a facet grid graph to visualise against elevation 
# Convert the "site" column to character in mastersheet_raw and update the dataset
mastersheet_raw <- mastersheet_raw %>%
  mutate(site = as.character(site))

# Perform the left join to include elevation in the zooplankton data sheet 
zooplankton2 <- zooplankton1 %>%
  left_join(select(mastersheet_raw, elevation_m, site), by = c("Site" = "site"))

#arrange by elevation 
zooplankton2 <- zooplankton2 %>%
  arrange(elevation_m)

#plotted abundance by elevation 
plot_1 <- ggplot(zooplankton2, aes(x = elevation_m, y = Count, fill = Order)) +
  geom_bar(stat = "identity", position = "dodge", width = 30) +
  labs(x = "Elevation (m)", y = "Abundance", fill = "Order") +
  facet_wrap(~ Survey_period, ncol = 1, scales = "free_y")

plot_1

#facet grid plot by order of zooplankton 
plot_2 <- ggplot(zooplankton2, aes(x = elevation_m, y = Count, fill = Order)) +
  geom_bar(stat = "identity", position = "dodge", width = 50) +
  scale_fill_brewer(palette = "Paired") +
  labs(x = "Elevation (m)", y = "Abundance", fill = "Order", title = "Zooplankton Abundance by Elevation and Order") +
  theme(plot.title = element_text(hjust = 0.5),
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  facet_grid(. ~ Order)

plot_2

```
