---
title: "Verreauxii_tadpole_code"
output: html_document
date: '2023-10-27'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Statistical analysis: 

(1) that Bd prevalence in tadpoles would be lower than in adults from the same breeding sites 

```{r}
install.packages("pacman")
pacman::p_load(DHARMa, lme4, rptR, ggplot2, dplyr, readr, tidyverse, patchwork, epitools, cowplot, here)

data <- read_csv(here("data", "clean_data.csv")) #all data 
zoop <- read_csv(here("data","zoop.csv")) #microfauna analysis summary
temp_summary <- read_csv(here("data","temp_summary.csv")) #ibutton data 

# Create a subset of data where record_type is equal to a specific value, e.g., "A"
tadpole_data <- subset(data, record_type == "Tadpole")
adult_data <- subset(data, record_type == "Adult")

```

```{r}

#tadpole only model 
glmm1 <- glmer(Bd_binary ~  scale(length_mm) + scale(mean_temp) + scale(surface_temp) + scale(raw_count) + scale(species_richness) + (1 | site), family = binomial, data = tadpole_data)
summarymodel <- summary(glmm1)
summarymodel

confint(glmm1)

simulationOutput <- simulateResiduals(fittedModel= glmm1)
simulationOutput
plot(simulationOutput)

##testing repeatability
rep1 <- rptProportion(Bd_binary ~  scale(length_mm) + scale(mean_temp) + scale(surface_temp) + scale(raw_count) + scale(species_richness) + (1 | site), grname=c("site"), data = tadpole_data, link = c("logit"))

```

Adult Model Only

```{r}

#adult only model 

glmm2 <- glmer(Bd_binary ~ scale(amb_temp) + scale(SVL_mm) + scale(elevation.x) +  (1 | site), family = binomial, data = adult_data)
model2 <- summary(glmm2)
model2$coefficients

confint(glmm2)

simulationOutput2 <- simulateResiduals(fittedModel= glmm2)
simulationOutput2

plot(simulationOutput2)

##testing repeatability
rep2 <- rptProportion(Bd_binary ~ amb_temp + SVL_mm + elevation.x +  (1 | site), data = adult_data, link = c("logit"))

```

```{r}
#all record types GLMM 
#issue with levels and record_type- need to check the levels before running 
str(data)
data$survey_period <- as.factor(data$survey_period)
data$survey_period <- as.factor(data$survey_period)
data$record_type <- factor(data$record_type, levels = unique(data$record_type))
levels(data$record_type)

glmm3 <- glmer(Bd_binary ~ record_type + survey_period + scale(amb_temp) + scale(elevation.x) + (1 | site), family = binomial, data = data)
model3 <- summary(glmm3)
model3$coefficients

confint(glmm3)

simulationOutput3 <- simulateResiduals(fittedModel= glmm3)
simulationOutput3

plot(simulationOutput3)

##elevation vs Bd binary- significant 
ggplot(data, aes(x = elevation.x, y = Bd_binary))+
  geom_point() + 
  geom_smooth()
  
rep3 <- rptProportion(Bd_binary ~ record_type + survey_period + amb_temp + elevation.x + (1 | site), data = data, link = c("logit"))

```

GLM for body size vs prevalence for tadpoles

```{r}
# Fit a GLM with binomial family to look at tadpole body size effects 

glm_model_body <- glm(Bd_binary ~ length_mm, family = poisson, data = all_data)

# Summarize the model
summary(glm_model_body)

```

##GLM's for microfauna 

(3) there would be a negative relationship between Bd prevalence in tadpoles and the abundance and diversity of microfauna, which can consume Bd zoospores.  
WAS found in the GLMM but not accurate yet. 

```{r}
#load zoop dataframe
zoop <- read_csv("zoop.csv")

##GLM to test abundance and diversity vs elevation 
glm_diversity <- glm(shannon_index ~ elevation_m, data = zoop, family = gaussian())
summary(glm_diversity)
```

```{r}
glm_abundance <- glm(count ~ elevation_m, data = zoop, family = poisson())
summary(glm_abundance)

```

### VISUALISATIONS ###

```
# Aggregate the data based on elevation, site, record_type, binary_bd, and Survey_period

aggregated_df <- data %>%
  group_by(elevation.x, site, record_type, survey_period, Bd_binary) %>%
  summarise(count = n(), .groups='drop')

# Calculate the total count per group (elevation, record_type, Survey_period)
  
total_count_per_group <- aggregated_df %>%
  group_by(record_type, site, survey_period) %>%
  summarise(total_count = sum(count)
  ) %>%
  ungroup()

# Include the total count back into the aggregated DataFrame
aggregated_df <- left_join(aggregated_df, total_count_per_group, by = c("site", "record_type", "survey_period"))

# Calculate the proportion for each group, Adults, TE, TL
aggregated_df$proportion <- aggregated_df$count / aggregated_df$total_count

# Create a new column with custom facet labels
aggregated_df$custom_facet <- case_when(
  aggregated_df$record_type == "Adult" & aggregated_df$survey_period == 1 ~ "Adult",
  aggregated_df$record_type == "Tadpole" & aggregated_df$survey_period == 1 ~ "Early Tadpole Sampling",
  aggregated_df$record_type == "Tadpole" & aggregated_df$survey_period == 2 ~ "Late Tadpole Sampling",
  TRUE ~ as.character(interaction(aggregated_df$record_type, aggregated_df$survey_period, lex.order = TRUE))
  )

# Convert elevation_m to numeric and sort it
aggregated_df$elevation.x <- as.numeric(as.character(aggregated_df$elevation.x))
aggregated_df <- aggregated_df[order(aggregated_df$elevation.x), ]

# Convert it back to factor with ordered levels
aggregated_df$elevation.x <- factor(aggregated_df$elevation.x, levels = unique(aggregated_df$elevation.x))
  
# Create the plot with site labels on the x-axis and adjusted bar width
```

PREVELANCE PLOT

```
prevalence_plot <- ggplot(aggregated_df, aes(x = as.factor(site), y = proportion, fill = as.factor(Bd_binary))) +
  geom_bar(stat = "identity", position = "fill", width = 0.7) +
  geom_text(aes(label = total_count, y = 1.02), vjust = 0.1, size = 3) +  # Add count label at the top of bars
  facet_wrap(~custom_facet, scales = "free_x") +
  scale_fill_manual(values = c("#56B4E9", "#E69F00"), labels = c("Negative", "Positive")) +
  labs(
    x = "Site",
    y = "Bd Prevalence Proportion",
    fill = "Bd Status"
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 70, hjust = 1))

prevalence_plot


```

(2) that Bd prevalence in tadpoles would decrease as the breeding season progressed and water temperatures increased, potentially creating warm microhabitat conditions where tadpoles could avoid or clear Bd infections. 
GLMM model meant to be the test here... but working that one out

- made overall plot of temps from ibuttons 
- comparison boxplot of time periods 

```
#line plot of temperature over time, grouped by site and coloured by depth category
ggplot(extracted_data, aes(x = date, y = temperature, color = depth_category)) +
  geom_line() +
  labs(x = "Date", y = "Temperature (°C)", color = "Depth Category") +
  facet_wrap(~ site, scales = "free_y", ncol = 2) +  # Facet by site, adjust the number of columns as needed
  theme_minimal()

#make a clear distinction between time periods 

# Calculate a reference date as 4 weeks (28 days) before the maximum date
reference_date <- max(extracted_data$date) - 28*24*3600  # 28 days in seconds

# Create a new column 'time_period' to categorize the dates
extracted_data <- extracted_data %>%
  mutate(time_period = ifelse(date >= reference_date, "72 Hours Before Collection", "72 Hours After Deployment"))

# Create side-by-side boxplots of temperature by elevation category, colored by time period
time_period_plot <- ggplot(extracted_data, aes(x = factor(elevation), y = temperature, fill = factor(time_period))) +
  geom_boxplot(position = "dodge") +
  labs(x = "Elevation", y = "Temperature (°C)", fill = "Time Period") +
  theme_minimal()

time_period_plot

```
- box plot shows clear increase in temperature. good way to visualise this? 


TADPOLES AND TEMPS 
- visualise amount of time spent in optimum Bd temperatures for the tadpoles  

```
##tadpole records 

#how many hours are the tadpoles in unfavourable conditions? 
ibutton <- ibutton_data_raw

df_unfavorable <- ibutton %>%
  group_by(site, date) %>%
  mutate(unfavorable_hours = sum(temperature > 25))

merged_data <- merge(ibutton, df_unfavorable, by = c("site", "date"))

#this calculates the number of unfavourable hours at each pond. 
unfavorable_hours_by_pond <- merged_data %>%
  group_by(site) %>%
  summarise(total_unfavorable_hours = sum(unfavorable_hours))

#need to change site to factor, not numeric
ggplot(data = unfavorable_hours_by_pond, aes(x = factor(site), y = total_unfavorable_hours)) +
  geom_bar(stat = "identity") +
  labs(x = "Site", y = "Total unfavorable hours") +
  ggtitle("Total Unfavorable Hours above 25 degrees Celcius") +
  theme(plot.title = element_text(hjust = 0.5)) +
  theme(axis.title = element_text(size = 12), axis.text = element_text(size = 10))

#now for under 4 degrees 
low_unfavorable <- ibutton %>%
  group_by(site, date) %>%
  mutate(unfavorable_hours = sum(temperature < 17))

merged_data2 <- merge(ibutton, low_unfavorable, by = c("site", "date"))

#this calculates the number of unfavourable hours at each pond. 
low_unfavorable_hours_by_pond <- merged_data2 %>%
  group_by(site) %>%
  summarise(total_unfavorable_hours = sum(unfavorable_hours))

#visualise this 
#need to change site to factor, not numeric

ggplot(data = low_unfavorable_hours_by_pond, aes(x = factor(site), y = total_unfavorable_hours)) +
  geom_bar(stat = "identity") +
  labs(x = "Site", y = "Total unfavorable hours") +
  ggtitle("Total Unfavorable Hours below 17 degrees Celcius") +
  theme(plot.title = element_text(hjust = 0.5)) +
  theme(axis.title = element_text(size = 12), axis.text = element_text(size = 10))

## daily min and max temps experienced by tadpoles 
# Read in the iButton data
ibutton_data <- ibutton

# Convert the date/time variable to a POSIXct format
ibutton_data$date <- as.POSIXct(ibutton_data$date, format = "%m/%d/%Y %H:%M")

# Create separate columns for date and time
ibutton_data$date <- as.Date(ibutton_data$date)
ibutton_data$time <- format(ibutton_data$date, "%H:%M:%S")

# Create separate columns for shallow and deep temperatures
ibutton_data1 <- pivot_wider(data = ibutton_data, names_from = site, values_from = temperature)

##too many duplicates- try mean 
# Calculate the daily mean temperatures for each site
daily_mean_temp <- ibutton_data %>%
  group_by(date, site, elevation) %>%
  summarize(mean_temp = mean(temperature, na.rm = TRUE))

# Convert pond to a factor
daily_mean_temp$site <- factor(daily_mean_temp$site)

#plot mean temp 
ggplot(data = daily_mean_temp, aes(x = date, y = mean_temp, color = site)) +
  geom_line() +
  scale_color_discrete(name = "Site") +
  labs(x = "Date", y = "Temperature (°C)") +
  ggtitle("Daily Mean Temperatures by Site") +
  theme(plot.title = element_text(hjust = 0.5), legend.position = "bottom") +
  theme(axis.title = element_text(size = 12), axis.text = element_text(size = 10))
  
#Calculate the daily min and max temperatures for each site
daily_temp_summary <- ibutton_data %>%
  group_by(date, site, elevation) %>%
  summarize(min_temp = min(temperature),
            max_temp = max(temperature))
            
# Convert elevation to a factor
daily_temp_summary$elevation <- factor(daily_temp_summary$elevation)

# Plot the daily min and max temperatures for each site
ggplot(data = daily_temp_summary, aes(x = date, y = max_temp, color = elevation)) +
  geom_line() +
  geom_line(aes(y = min_temp)) +
  scale_color_discrete(name = "Elevation") +
  labs(x = "Date", y = "Temperature (°C)") +
  ggtitle("Daily Min and Max Temperatures by Elevation") +
  theme(plot.title = element_text(hjust = 0.5), legend.position = "bottom") +
  theme(axis.title = element_text(size = 12), axis.text = element_text(size = 10))

##remove the outliers if needed
# Filter out rows with temperatures above 35 degrees
daily_temp_summary_filtered <- daily_temp_summary %>%
  filter(max_temp <= 35, min_temp <= 35)

# Convert elevation to a factor
daily_mean_temp$elevation <- factor(daily_mean_temp$elevation)

# daily mean ggplot
dailymeantemp_plot <- ggplot() +
  geom_line(data = daily_mean_temp, aes(x = date, y = mean_temp, color = elevation)) +
  scale_color_discrete(name = "Site") +
  labs(x = "Date", y = "Temperature (°C)") +
  ggtitle("Daily Mean Temperatures (°C) Over Sampling Period") +
  geom_hline(yintercept = 25, color = "red", linetype = 2, linewidth = 1.0) +
  geom_hline(yintercept = 17, color = "red", linetype = 2, linewidth = 1.0) +
  theme(plot.title = element_text(hjust = 0.5), legend.position = "right") + theme(axis.title = element_text(size = 12), axis.text = element_text(size = 10))

dailymeantemp_plot

##plot the daily min and max temps, with chytrid limits - GREAT FOR PAPER 
#Plot the daily min and max temperatures for each site
minmax_plot <- ggplot(data = daily_temp_summary_filtered, aes(x = date, y = max_temp, color = site)) +
  geom_line(aes(x= date, y = max)) +
  scale_color_discrete(name = "Site") +
  labs(x = "Date", y = "Temperature (°C)") +
  ggtitle("Daily Min and Max Temperatures by Site") +
  geom_hline(yintercept = 25, color = "red", linetype = 2, linewidth = 1.0) +
  geom_hline(yintercept = 4, color = "red", linetype = 2, linewidth = 1.0) +
  theme(plot.title = element_text(hjust = 0.5), legend.position = "bottom") + theme(axis.title = element_text(size = 12), axis.text = element_text(size = 10))

# Filter Bd_records for records with record_type "Tadpole"
tadpole_records <- Bd_records %>%
  filter(record_type == "Tadpole", !is.na(ID))

# Now create the plot using the cleaned data
tadpole_plot <- ggplot(tadpole_records, aes(x = date, y = surface_temp, color = Bd)) +
geom_point()+
  geom_hline(yintercept = 25, color = "red", linetype = 2, linewidth = 1.5) +
  geom_hline(yintercept = 17, color = "red", linetype = 2, linewidth = 1.5) +
  labs(x = "Date", y = "Temperature (°C)") +
  ggtitle("Temperatures of Tadpoles at Time of Capture")

tadpole_plot
```

FINAL TEMP PLOT 

```
# Combine the two plots
library(cowplot)

# Combine plots
# Convert date values to Date format
dailymeantemp_plot$data$date <- as.Date(dailymeantemp_plot$data$date)
tadpole_plot$data$date <- as.Date(tadpole_plot$data$date)

# Set the x-axis limits based on the date range of your data
date_range <- range(dailymeantemp_plot$data$date, tadpole_plot$data$date)

# Create the plots
dailymeantemp_plot <- dailymeantemp_plot +
  xlim(date_range)

tadpole_plot <- tadpole_plot +
  xlim(date_range)

# Combine plots side by side
combined_plot <- plot_grid(dailymeantemp_plot, tadpole_plot, labels = c("A", "B"), ncol = 1)

# Display the combined plot
print(combined_plot)

```

Micropredator data: 

```
# Create a bar plot of Shannon index by Elevation and Survey_period
ggplot(zoop_summary_elevation, aes(x = factor(elevation_m), y = mean_shannon, fill = factor(Survey_period))) +
  geom_bar(stat = "identity", position = "dodge", width = 0.7) +
  labs(x = "Elevation (m)", y = "Mean Shannon Index", fill = "Survey Period") +
  facet_wrap(~ Survey_period, ncol = 1, scales = "free_y") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1)) + #rotates the labels 
  scale_x_discrete(labels = function(x) format(as.numeric(x), scientific = FALSE, trim = TRUE)) #reduces the decimal places of elevation
  
#export a table 
install.packages("gt")
library(gt)

zoop_summary %>%
  gt() %>%
  tab_header( title = "Zooplankton Species Richness and Shannon Index Summary")

  
##zooplankton abundance ####
# simplify the columns 
zooplankton1 <- zooplankton %>%
    mutate(Site = as.character(Site)) %>%
  select(c(Family, Order, Site, Survey_period, Count))
 
#order across all sites, in two sampling periods 
ggplot(zooplankton1, aes(x = Site, y = Count, fill = Order)) +
  geom_bar(stat = "identity", position = "dodge", width = 0.7) +
  labs(x = "Site", y = "Abundance", fill = "Order") +
  scale_x_discrete(limits = unique(zooplankton1$Site)) +
  facet_wrap(~ Survey_period, ncol = 1, scales = "free_y") +
  theme(axis.text.y = element_text(angle = 90, hjust = 1))

##making a facet grid graph to visualise against elevation 
# Convert the "site" column to character in mastersheet_raw and update the dataset
mastersheet_raw <- mastersheet_raw %>%
  mutate(site = as.character(site))

# Perform the left join to include elevation in the zooplankton data sheet 
zooplankton2 <- zooplankton1 %>%
  left_join(select(mastersheet_raw, elevation_m, site), by = c("Site" = "site"))

#arrange by elevation 
zooplankton2 <- zooplankton2 %>%
  arrange(elevation_m)

#plotted abundance by elevation 
plot_1 <- ggplot(zooplankton2, aes(x = elevation_m, y = Count, fill = Order)) +
  geom_bar(stat = "identity", position = "dodge", width = 30) +
  labs(x = "Elevation (m)", y = "Abundance", fill = "Order") +
  facet_wrap(~ Survey_period, ncol = 1, scales = "free_y")

plot_1

#facet grid plot by order of zooplankton 
plot_2 <- ggplot(zooplankton2, aes(x = elevation_m, y = Count, fill = Order)) +
  geom_bar(stat = "identity", position = "dodge", width = 50) +
  scale_fill_brewer(palette = "Paired") +
  labs(x = "Elevation (m)", y = "Abundance", fill = "Order", title = "Zooplankton Abundance by Elevation and Order") +
  theme(plot.title = element_text(hjust = 0.5),
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  facet_grid(. ~ Order)

plot_2

```

Making Map of sites 

```{r}
# Load necessary libraries

library(dplyr)
library(sf)
library(ggplot2)
library(dplyr)
library(raster)
library(rasterVis)  # Optional, for advanced raster visualization

# Read the data
gps_data <- read.csv("~/Library/CloudStorage/OneDrive-AustralianNationalUniversity/Jordann PhD Files/whistling tree frog/whistling_tree_frog/combined_visits.csv")

# Determine the type of each site based on 'visit_type' and 'site'
site_types <- gps_data %>%
  group_by(site) %>%
  summarize(visit_type_combined = paste(unique(visit_type), collapse = ",")) %>%
  mutate(site_type = case_when(
    grepl("Tadpole", visit_type_combined) & grepl("Adult", visit_type_combined) ~ "Both",
    grepl("Tadpole", visit_type_combined) ~ "Tadpole",
    grepl("Adult", visit_type_combined) ~ "Adult"
  ))

# Join this information back to the gps_data
gps_data <- gps_data %>%
  left_join(site_types, by = "site")

#Transform GPS data to spatial data
gps_points <- st_as_sf(gps_data, coords = c("lon", "lat"), crs = 4326)

# Load Australia SA2 boundary data
australia_sa2_shape <- st_read('~/Library/CloudStorage/OneDrive-AustralianNationalUniversity/Jordann PhD Files/whistling tree frog/whistling_tree_frog/SA2_2021_AUST_SHP_GDA2020')

# Filter to get only the ACT regions
act_regions <- australia_sa2_shape %>% 
  filter(STE_NAME21 == "Australian Capital Territory")

# Dissolve all ACT regions into a single boundary
act_boundary <- act_regions %>%
  summarize(geometry = st_union(geometry)) %>%
  st_cast("POLYGON")

elevation_data <- raster("~/OneDrive - Australian National University/Jordann PhD Files/whistling tree frog/whistling_tree_frog/5m_DEM.tif")

# Check CRS of all layers
crs_act_boundary <- st_crs(act_boundary)
crs_gps_points <- st_crs(gps_points)
crs_elevation_data <- crs(elevation_data)

# Print CRS for verification
print(crs_act_boundary)
print(crs_gps_points)
print(crs_elevation_data)

# Assuming you want to transform everything to the CRS of the elevation data
# Transform ACT boundary and GPS points to match elevation data CRS
act_boundary_transformed <- st_transform(act_boundary, crs_elevation_data)
gps_points_transformed <- st_transform(gps_points, crs_elevation_data)

# Filter to include only sites with both tadpoles and adults
gps_points_filtered <- gps_points_transformed %>% 
  filter(site_type == "Both")

# Define extent for ACT
extent_of_interest <- extent(c(148.7, 149.4, -36, -35.1))

# Clip raster to extent of interest
elev_raster_clipped <- crop(elevation_data, act_boundary_transformed)

# Define the aggregation factor
aggregation_factor <- 5

# Aggregate the raster
elev_raster_aggregated <- aggregate(elev_raster_clipped, fact = aggregation_factor, fun = mean)

# Ensure that the elevation data is in the same CRS as the ACT boundary
elevation_data_transformed <- projectRaster(elev_raster_aggregated, crs = crs(act_boundary))

# Perform the masking operation
elevation_data_masked <- mask(elevation_data_transformed, act_boundary)

# Convert the masked raster to a dataframe for ggplot2
elevation_df <- raster::rasterToPoints(elevation_data_masked)
elevation_df <- data.frame(elevation_df)

### final map #### 
main_map <- ggplot() +
  geom_raster(data = elevation_df, aes(x = x, y = y, fill = X5m_DEM), alpha = 0.7) +
  scale_fill_viridis_c() +
  geom_sf(data = act_boundary, fill = NA, color = "black") +
  geom_sf(data = gps_points_filtered, color = "black", shape = 21, fill = "white", size = 2, stroke = 0.5) +
  labs(fill = "Elevation", 
       color = "Site Type",
       shape = "Site Type") +
  theme_void() +
coord_sf(crs = crs_elevation_data, xlim = c(min(st_coordinates(act_boundary)[,1]), max(st_coordinates(act_boundary)[,1])), ylim = c(min(st_coordinates(act_boundary)[,2]), max(st_coordinates(act_boundary)[,2])))

map <- main_map + 
  theme(legend.position = "left")

### additions to the map ###
### inlay map 
library(ggplot2)
library(sf)
library(rnaturalearth)

world <- ne_countries(scale = "medium", returnclass = "sf")

# Extract Australia
australia <- world[world$admin == "Australia", ]

##check projection 
crs_australia <- st_crs(australia)
crs_act_boundary <- st_crs(act_boundary)

##transform CRS
australia_transformed <- st_transform(australia, crs = crs_act_boundary)

# Print the updated map
australia_act_map <- ggplot() +
  geom_sf(data = australia_transformed, fill = NA, color = "black") +
  geom_sf(data = act_boundary_transformed, fill = "blue", color = "black") +
  theme_void()

australia_act_map

##bounding box to show ACT 
australia_map <- australia_act_map +
  geom_rect(
    aes(
      xmin = 146, xmax = 151, 
      ymin = -38, ymax = -34
    ),
    color = "black", fill = NA, size = 1
  ) +
  theme_void()

australia_map

# Use cowplot to add the inset map
library(cowplot)

final_map_figure <- ggdraw() +
  draw_plot(map) +
  draw_plot(australia_map, x = 0.60, y = 0.01, width = 0.4, height = 0.4)

print(final_map_figure)

ggsave("~/OneDrive - Australian National University/Jordann PhD Files/whistling tree frog/whistling_tree_frog/final_map.png", final_map_figure, width = 10, height = 8, units = "in", dpi = 300)

```

